Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes

### 2022-11-14 Manuscript.tex
fixing bibtex entry

\nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
\nomenclature{$\hat{\bf n}_i$}{Reference Orientation i}

@@

Using our simulation from Section~\ref{sec:2.1} we calculated the
instantaneous signal from our dimer at each time step using MSTM.  We
applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:2.2} ---
and the reference orientation with the highest probability is taken as
our estimation of the dimer's instantaneous orientation.  The prior of
the reference orientations $p_i(\hat{\bf n})$ is defined at each time
step by accounting for the physical distance between our previous
estimation $\hat{\bf n}_{i, \ t - \Delta t}$ and each reference
orientation $\hat{\bf n}$
\begin{align}
  r_i(t, \hat{\bf n}_i)
  &= \frac{1}{2}((\hat{\bf n}_{i,\ t} \cdot \hat{\bf n}_{t-\Delta t})-1)
  \\
  p_i(\hat{\bf n})
  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
{\sum_{j=1}^{30}e^{\beta(r_j(t, \hat{\bf n}_i)-1)}}
\end{align}
where $\beta$ is a weighting factor that controls the significance of
our priori estimate.  The simulation data from Section~\ref{sec:2.1}
was used to check if the estimation was correct by comparing the
estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation
closest to the instantaneous orientation, denoted as
$\hat{\bf n}_{best}$.  To measure the error in the model's estimate
the distribution's of $p(\hat{\bf n}| y(\bar{\bf s}))$ and an
ideal result, $p_ideal$, were compared.  An ideal distribution is one
that is 0 for every $\hat{\bf n}$ apart from the correct result.
\begin{align}
p_{ideal} = 
\begin{cases}
	1 & \text{when $\hat{\bf n}_i$ = $\hat{\bf n}_{best}$}\\
	0 & \text{anywhere else}
\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation.  The model's error was
measured by calculating the Kullback-Leibler divergence $K_l$ between
the two distributions
\begin{align}
  K_{l, \#}(p_i(\hat{\bf n}| y_k(\bar{\bf s})) &\parallel p_{ideal}) =\nonumber
  \\
&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y(\bar{\bf s}))}
\right]
\end{align}
where a more negative value of $K_l$ indicates a poorer estimation of
the dimer's instantaneous orientation.  By summing the divergence of
each measurement across the entire simulation we get an evaluation of
how well the model performed in estimating the dimer's orientation.
To compare between different iterations of our model the we compare
our result of $K_l$ to a worst case scenario and evaluate how much it
improves upon this, denoted as $F(K_l)$:
\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

The signal error $\epsilon$ and the weighting factor $\beta$ are known
quantities whose influence can be measured directly; however, our
choice of angles and how they correlate with $F(K_l)$ is not easily
understood Evaluating every possible configuration manually is
inefficient, therefore we utilised Ultra-nest to apply a Monte Carlo
simulation to our model. Ultra-nest is a sophisticated python package
for fitting models with complex parameter correlations
\cite{Buchner2016Ultranest}.  UltraNest implements a variation of
Monte Carlo integration, nested-sampling, where a likelihood contour
is used to update a group of live points chosen from the prior
estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

The Brownian OT software was used to simulate the motion of a trapped
dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first $10$\,seconds of
entering the optical trap.  The initial orientation was assumed as
strictly vertical (in line with the beam propagation direction). The
dimer's position orientation was recorded every $10\,{\rm \mu s}$ for
using as a test dataset for our model.

@@

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a
full $180^{\circ}$ rotation upon entering the trap. Typically
horizontal alignment of a dimer is unstable and will result in the
particle rotating to align along its vertical axis. It is interesting
to note that the dimer is furthest from the trap centre as it goes
into a horizontal orientation before drawing closer again as it
inverts completely. Further simulations of differently sized dimers
showed similar results, but only when $a_1 \geq 2a_2$. Dimers with
more symmetrical size ratios immediately aligned into a fixed vertical
position.
%
In Vigilante's work with dimers \cite{Vigilante2020Brownian_OT}
simulations of trapped symmetrical dimers was investigated; their
findings showed that the optical torque on the dimer goes to zero
while aligned vertically and is at its maximum in a horizontal
alignment.  Therefore, the inversion of an asymmetric dimer suggests
that if the size difference is significant the optical torque goes is
minimal for a dimer in both horizontal and vertical orientations. Once
we have our experimental set up complete we can confirm this by
trapping an asymmetric dimer and monitoring its orientation. The data
from \figurename{ 2a} was used to test the efficacy of our predictive
model. To visualise how well our model tracks the dimer's orientation
we plot the radial distance between our model's prediction and the
true orientation, and between the best possible orientation and the
true orientation. For a perfect estimation these two plots should
overlap, while we cannot achieve an exact estimation we can optimise
it by adjusting the expected signal error and weighting factor.

@@

\subsection{Parameter Optimisation Results}
\label{sec:3.3}
We wanted to see if our choice of angles plays a significant role in our model's efficacy. We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$. We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.

@@

Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer. Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle. While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we can't achieve any greater accuracy than dividing the orientation space into 30 reference orientations. We hope to develop our model further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.

>>>>

\nomenclature{$\bar{\bf s}$}{Dimer's instantaneous orientation}
\nomenclature{$\hat{\bf n}_\alpha$}{Reference Orientation $\alpha$}

@@

Using our simulation from Section~\ref{sec:2.1} we calculated the
instantaneous signal from our dimer at each time step using MSTM.  We
applied Eq.~\eqref{eq:bayes} --- covered in Section~\ref{sec:2.2} ---
and the reference orientation with the highest probability is taken as
our estimation of the dimer's instantaneous orientation.  The prior of
the reference orientations $p(\hat{\bf n}_\alpha)$ is defined at each
time step by accounting for the physical distance between the previous
estimate $\hat{\bf n}_{\alpha}(t - \Delta t)$ and each reference
orientation $\hat{\bf n}_\alpha$
\begin{align}
  r_i(t, \hat{\bf n}_\alpha)
  &= \frac{1}{2}((\hat{\bf n}_{\alpha}(t) \cdot \hat{\bf n}_{t-\Delta t})-1)
  \\
  p_i(\hat{\bf n})
  &= \frac{e^{\beta (r_i(t,\hat{\bf n}_i)-1)}}
{\sum_{j=1}^{n_{\rm ref}}e^{\beta(r_j(t, \hat{\bf n}_\alpha)-1)}}
\end{align}
where $\beta$ is a weighting factor that controls the significance of
our priori estimate.  The simulation data from Section~\ref{sec:2.1}
was used to check if the estimation was correct by comparing the
estimate provided by Eq.~\eqref{eq:bayes} to the reference orientation
closest to the instantaneous orientation, denoted as
$\hat{\bf n}_{best}$.  To measure the error in the model's estimate
the distribution's of $p(\hat{\bf n}| y(\bar{\bf s}))$ and an
ideal result, $p_ideal$, were compared.  An ideal distribution is one
that is 0 for every $\hat{\bf n}$ apart from the correct result.
\begin{align}
p_{ideal} = 
\begin{cases}
	1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
	0 & \text{anywhere else}
\end{cases}
\end{align}
The distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation.  The model's error was
measured by calculating the Kullback-Leibler divergence $K_l$ between
the two distributions
\begin{align}
  K_{l, \#}(p(\hat{\bf n}_\alpha| y_1, y_2, y_3), p_{ideal})
  =\nonumber
  \\
&p_{ideal}\ln \left[\frac{p_{ideal}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
\right]
\end{align}
where a more negative value of $K_l$ indicates a poorer estimation of
the dimer's instantaneous orientation.  By summing the divergence of
each measurement across the entire simulation we get an evaluation of
how well the model performed in estimating the dimer's orientation.
To compare between different iterations of our model the we compare
our result of $K_l$ to a worst case scenario and evaluate how much it
improves upon this, denoted as $F(K_l)$:
\begin{align}
K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#} \\
K_{l, \ worst} &= -\sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/30} \right] \\
F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

@@

The signal error $\epsilon$ and the weighting factor $\beta$ are known
quantities whose influence can be measured directly; however, our
choice of angles and how they correlate with $F(K_l)$ is not easily
understood Evaluating every possible configuration manually is
inefficient, therefore we utilised UltraNest to apply a Monte Carlo
simulation to our model.  UltraNest is a Python package for fitting
models with complex parameter correlations
\cite{Buchner2016Ultranest}.  UltraNest implements a variation of
Monte Carlo integration, nested-sampling, where a likelihood contour
is used to update a group of live points chosen from the prior
estimation ($0^{\circ} \leq \theta_k \leq 180^{\circ}$).

@@

The Brownian OT software was used to simulate the motion of a trapped
dimer ($a_1=1\,{\rm \mu m}$, $a_2=0.5\,{\rm \mu m}$) over the first
$10$\,seconds of entering the optical trap.  The initial orientation
was assumed as strictly vertical (in line with the beam propagation
direction). The dimer's position orientation was recorded every
$10\,{\rm \mu s}$ for using as a test dataset for our model.

@@

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a
full $180^{\circ}$ rotation upon entering the trap.  Typically
horizontal alignment of a dimer is unstable and will result in the
particle rotating to align along its vertical axis.  It is interesting
to note that the dimer is furthest from the trap centre as it goes
into a horizontal orientation before drawing closer again as it
inverts completely. Further simulations of differently sized dimers
showed similar results, but only when $a_1 \geq 2a_2$.  Dimers with
more symmetrical size ratios immediately aligned into a fixed vertical
position.
%
In the simulations of Vigilante~\emph{et~al}.\
\cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were
investigated; their findings showed that the optical torque on the
dimer goes to zero while aligned vertically and is at its maximum in a
horizontal alignment.  Therefore, the inversion of an asymmetric dimer
suggests that if the size difference is significant the optical torque
goes is minimal for a dimer in both horizontal and vertical
orientations. Once we have our experimental set up complete we can
confirm this by trapping an asymmetric dimer and monitoring its
orientation.  The data from \figurename{ 2a} was used to test the
efficacy of our predictive model.  To visualise how well our model
tracks the dimer's orientation we plot the radial distance between our
model's prediction and the true orientation, and between the best
possible orientation and the true orientation.  For a perfect
estimation these two plots should overlap, while we cannot achieve an
exact estimation we can optimise it by adjusting the expected signal
error and weighting factor.

@@

\subsection{Parameter Optimisation Results}
\label{sec:3.3}

We wanted to see if our choice of angles plays a significant role in
our model's efficacy. We employed ultranest to analyse the parameter
space for our three angles by sampling across the priori of
$10^{\circ} \ \& \ 180^{\circ}$.   We kept both $\beta$ and $\sigma$
constant during this so we could as these are parameters that can be
changed more freely.

@@

Figure~\ref{fig:corner} is the resulting corner plot from our
ultranest sampler, darker areas indicate a choice of angles where the
model performed better where as lighter areas are where the model did
poorer.  Overall from the corner plot provided, it is clear that there
is no direct correlation between our method's performance and the
choice of detection angle. While the choice of angles does not provide
a direct improvement in our estimation it does determine how finely we
can estimate the orientation, for choices of angles that are close
together the minimum spacing between our signals is so small that we
cannot achieve any greater accuracy than dividing the orientation
space into $n_{\rm ref}=30$ reference orientations.  We hope to
develop our method further in the future by being able to determine
the maximum accuracy given that we know the signal error of our
measurements.
